---
title: "Basic example"
author: "Carlos Antonio Rodríguez-Saltos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we will report the daily key presses that a bird has made. The data will be analyzed only until one of the songs was exhausted each day.

## Importing data
First, we will import our data. To do so, we need only the name of the directory containing the files to be imported.

`songPreference` already comes with sample data from one bird, named ZF1536. We will import the data from this bird. For that, we need to know where the installation directory of `songPreference` resides.

```{r}
songPreference_dir <- find.package(package = "songPreference")
data_dir <- file.path(songPreference_dir, "ZF1536")
print(data_dir)
```

Next, we import the data using the command `loadOC` from `songPreference`. To use this command, we need to load `songPreference`.

```{r}
library(songPreference)
data <- loadOC(birDir = data_dir)
head(data)
```

## Filtering data
We will now filter the data to keep only records until one of the songs was exhausted, each day. This is done by using the function `filterByLastPb `.

```{r}
data_filtered <- filterByLastPb(data = data, filtype = "onesong")
```

## Getting the press count
A summary of daily presses for each key can be obtained using the function `getKeyCount`.

```{r}
daily_summary <- getKeyCount(data = data_filtered)
head(daily_summary)
```

The date is in numerical format. See `?Date` for more information.

## Plotting the proportion of daily presses
We will now plot the above data. We will build a plot of stacked areas representing the proportion of presses for each key along time.

First, we need to ccalculate the proportion of presses for key A.

```{r}
proportion_A <- apply(X = daily_summary, MARGIN= 1, FUN = function(x) {
  x[1]/(x[1] + x[2])
})
head(proportion_A)
```

Now we need to convert the dates in numerical format to a more legible format.

```{r}
names(proportion_A) <- as.Date(as.numeric(names(proportion_A)), 
                               origin= "1970-01-01")
head(proportion_A)
```

To facilitate the construction of the plots, we will arrange the data into a dataframe.
```{r}
proportions <- data.frame(date= as.Date(names(proportion_A)), 
                          A= proportion_A, 
                          B= 1 - proportion_A)
head(proportions)
```

The dataframe is in wide format. It must be converted to long format before plotting.

```{r}
library(reshape2)
proportions_long <- melt(data = proportions, id= "date", value.name = "proportion",
                         variable.name = "key")
head(proportions_long)
```

Now we can make the plot.

```{r fig.width=7}
library(ggplot2)
p <- ggplot(data = proportions_long, mapping = aes(x = date, y = proportion, fill= key))
p <- p + geom_area()
print(p)
```

## Exporting the data
Data can be exported through the use of a basic function such as `write.csv`. Check the help file for that function for more information. You can export either `data`, `daily_summary`, or `proportions_long` in this way.
