---
title: "Decide whether to run a reversal"
author: "Carlos Antonio Rodríguez-Saltos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We will analyze SingSparrow data to decide whether the key contingencies must be switched.

## Import data

We will assume that the last rows of the dataframe contain the most recent keypresses.

```{r}
library(songPreference)
data <- loadOC(birDir = file.path(find.package("songPreference"), "ZF1536"))
```

## Check whether to switch contingencies

The main function here is `shouldISwitch`. It will print out the decission of whether to switch or not.

```{r}
reversal_decision <- shouldISwitch(data= data, 
                                   alpha = 0.05, 
                                   consecutive_days = 5, 
                                   filtype= "onesong")
```

The first two columns are the daily press counts for A and B. The third column contains the p value from the statistical test of significant difference between the key press counts. The row numbers are a way of representing dates. They correspond to the number of days since 1970-01-01. This type of numeric representation is usually a convenient way for R to store information on dates. 

Finally, the last line contains the decision to run or not to run a reversal.
